---
title: "Decision Trees Mushrooms"
author: "Cesar A. Sanchez"
date: "7/12/2020"
input: https://www.kaggle.com/kopylovlvad/mushrooms-classification-with-decision-tree/data?select=mushrooms.csv
output: html_notebook
---

Mushrooms data set

Attribute Information: (classes: edible=e, poisonous=p)
cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s
cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s
cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
bruises: bruises=t,no=f
odor: almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s
gill-attachment: attached=a,descending=d,free=f,notched=n
gill-spacing: close=c,crowded=w,distant=d
gill-size: broad=b,narrow=n
gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y
stalk-shape: enlarging=e,tapering=t
stalk-root: bulbous=b,club=c,cup=u,equal=e,rhizomorphs=z,rooted=r,missing=?
stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s
stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
veil-type: partial=p,universal=u
veil-color: brown=n,orange=o,white=w,yellow=y
ring-number: none=n,one=o,two=t
ring-type: cobwebby=c,evanescent=e,flaring=f,large=l,none=n,pendant=p,sheathing=s,zone=z
spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y
population: abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y
habitat: grasses=g,leaves=l,meadows=m,paths=p,urban=u,waste=w,woods=d

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(ggplot2)
library(ggmosaic)
library (randomForest)
library(moderndive)
library(tidyverse)
library(patchwork)
library(tidymodels)
library(modeldata)
library(baguette)
library(ROCR)
library(randomForest)
library(JOUSBoost)
library(caret)
```

```{r}
mushroom <- read.csv("C:\\Users\\cesarasa\\Documents\\CASM stuff\\Maestria Ciencia de Datos ITESO\\Verano2020\\Arboles de regresion\\Lecture final project\\datasets_478_974_mushrooms.csv")
#str(mushroom)
```
cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s
cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s
cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
bruises: bruises=t,no=f
odor: almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s
gill-attachment: attached=a,descending=d,free=f,notched=n
gill-spacing: close=c,crowded=w,distant=d
gill-size: broad=b,narrow=n
gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y
stalk-shape: enlarging=e,tapering=t
stalk-root: bulbous=b,club=c,cup=u,equal=e,rhizomorphs=z,rooted=r,missing=?
stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s
stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o,pink=p,red=e,white=w,yellow=y
veil-type: partial=p,universal=u
veil-color: brown=n,orange=o,white=w,yellow=y
ring-number: none=n,one=o,two=t
ring-type: cobwebby=c,evanescent=e,flaring=f,large=l,none=n,pendant=p,sheathing=s,zone=z
spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y
population: abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y
habitat: grasses=g,leaves=l,meadows=m,paths=p,urban=u,waste=w,woods=d
```{r}
# mushroom$class <- as.character(mushroom$class)
mushroom$class[mushroom$class == 'p'] <- 'poisonous'
mushroom$class[mushroom$class == 'e'] <- 'edible'
# mushroom$cap.shape[mushroom$cap.shape == 'b'] <- 'bell'
# mushroom$cap.shape[mushroom$cap.shape == 'c'] <- 'conical'
# mushroom$cap.shape[mushroom$cap.shape == 'x'] <- 'convex'
# mushroom$cap.shape[mushroom$cap.shape == 'f'] <- 'flat'
# mushroom$cap.shape[mushroom$cap.shape == 'k'] <- 'knobbed'
# mushroom$cap.shape[mushroom$cap.shape == 's'] <- 'sunken'
# mushroom$cap.surface[mushroom$cap.surface == 'f'] <- 'fibrous'
# mushroom$cap.surface[mushroom$cap.surface == 'g'] <- 'grooves'
# mushroom$cap.surface[mushroom$cap.surface == 'y'] <- 'scaly'
# mushroom$cap.surface[mushroom$cap.surface == 's'] <- 'smooth'
# cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
# mushroom$cap.color[mushroom$cap.color == 'n'] <- 'brown'
# mushroom$cap.color[mushroom$cap.color == 'b'] <- 'buff'
# mushroom$cap.color[mushroom$cap.color == 'c'] <- 'cinnamon'
# mushroom$cap.color[mushroom$cap.color == 'g'] <- 'gray'
# mushroom$cap.color[mushroom$cap.color == 'r'] <- 'green'
# mushroom$cap.color[mushroom$cap.color == 'p'] <- 'pink'
# mushroom$cap.color[mushroom$cap.color == 'u'] <- 'purple'
# mushroom$cap.color[mushroom$cap.color == 'p'] <- 'pink'
```

```{r}
glimpse(mushroom)
```


```{r}
g <- ggplot(mushroom,aes(cap.shape))
g + geom_bar(aes(fill=class),width=0.5) +
  theme(axis.text.x = element_text(angle=0, vjust=0.7)) +
  labs(title="Mushroom edible and poison by cap shape",
       subtitle="cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s")
```

```{r}
g <- ggplot(mushroom,aes(cap.color))
g + geom_bar(aes(fill=class),width=0.5) +
  theme(axis.text.x = element_text(angle=0, vjust=0.7)) +
  labs(title="Mushroom edible and poison by Cap Color",
       subtitle="cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y")
```

```{r}
g <- ggplot(mushroom,aes(odor))
g + geom_bar(aes(fill=class), color='black',width=1.0) +
  theme(axis.text.x = element_text(angle=0, vjust=0.7)) +
  labs(title="Mushroom edible and poison by Odor",
       subtitle="odor: almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s")

#ggplot(mpg, aes(x = hwy)) + geom_histogram(aes(y=..density..), color = "grey30", fill = "blue",alpha=0.5) #+geom_density(alpha = .2, fill = "antiquewhite3") + labs(title="Bin width 1")
```


```{r}
g <- ggplot(mushroom,aes(population))
g + geom_bar(aes(fill=class), color='black', width=1) +
  theme(axis.text.x = element_text(angle=0, vjust=0.7)) +
  labs(title="Mushroom edible and poison by population of mushroom",
       subtitle="population: abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y")
```

```{r}
g <- ggplot(mushroom,aes(gill.color))
g + geom_bar(aes(fill=class), color='black', width=1.0) +
  theme(axis.text.x = element_text(angle=0, vjust=0.7)) +
  labs(title="Mushroom edible and poison by gill color",
       subtitle="black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y")
```


```{r}
ggplot(mushroom) + 
   geom_mosaic(aes(x = product(odor), fill=class)) +
  ggplot2::labs(x='odor',title='f(odor | class)', subtitle="x:almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s") + 
  facet_grid(cols = vars(odor))
  #facet_grid(rows = vars(odor), cols = vars(cap.color))
# + facet_grid(class~.)
```


```{r}
ggplot(mushroom) + 
   geom_mosaic(aes(x = product(cap.color), fill=class)) +
  ggplot2::labs(x='cap.color',title='f(cap.color| class)', subtitle="x:brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y") + 
  facet_grid(cols = vars(cap.color))
```

```{r}
ggplot(mushroom) + 
   geom_mosaic(aes(x = product(population), fill=class)) +
  ggplot2::labs(x='population',title='f(population | class)', subtitle="abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y") + 
  facet_grid(cols = vars(population))
# + facet_grid(class~.)
```


```{r}
ggplot(mushroom) + 
   geom_mosaic(aes(x = product(gill.color), fill=class)) +
  ggplot2::labs(x='gill.color',title='f(gill.color| class)', subtitle="x:black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y") + 
  facet_grid(cols = vars(gill.color))
```

```{r}
ggplot(mushroom) + 
   geom_mosaic(aes(x = product(cap.surface ), fill=class)) +
  ggplot2::labs(x='cap.surface',title='f(gill.color| class)', subtitle="x:fibrous=f,grooves=g,scaly=y,smooth=s") + 
  facet_grid(cols = vars(cap.surface ))
```

```{r}

ggplot(mushroom) + 
   geom_mosaic(aes(x = product(habitat), fill=class)) +
  ggplot2::labs(x='habitat',title='f(gill.color| class)', subtitle="x:grasses=g,leaves=l,meadows=m,paths=p,urban=u,waste=w,woods=d") + 
  facet_grid(cols = vars(habitat ))
```


```{r}
g1 <- ggplot(mushroom) + 
   geom_mosaic(aes(x = product(odor), fill=class)) +
  ggplot2::labs(x='odor',title='f(odor | class)', subtitle="x:almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s") + 
  facet_grid(cols = vars(odor))
  #facet_grid(rows = vars(odor), cols = vars(cap.color))
# + facet_grid(class~.)
g2 <- ggplot(mushroom) + 
   geom_mosaic(aes(x = product(cap.color), fill=class)) +
  ggplot2::labs(x='cap.color',title='f(cap.color| class)', subtitle="x:brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y") + 
  facet_grid(cols = vars(cap.color))

g3 <- ggplot(mushroom) + 
   geom_mosaic(aes(x = product(population), fill=class)) +
  ggplot2::labs(x='population',title='f(population | class)', subtitle="abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y") + 
  facet_grid(cols = vars(population))
# + facet_grid(class~.)
g4 <- ggplot(mushroom) + 
   geom_mosaic(aes(x = product(gill.color), fill=class)) +
  ggplot2::labs(x='gill.color',title='f(gill.color| class)', subtitle="x:black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y") + 
  facet_grid(cols = vars(gill.color))

#g1/g2
```

FUll exploration  with a contingence Table to evaluate if a mushroom is poisonous or edible depending the
feature assosiated
```{r}
mush_features <- colnames(mushroom)[-1]
table_res <- lapply(mush_features, function(x) {table(mushroom$class, mushroom[,x])})
names(table_res) <- mush_features
table_res
```


###########################################################################################
Insert decision Tree model for Random Forest to classify  edible or poisonous mushroom
###########################################################################################

```{r}
library(FeatureHashing)
library(xgboost)
mush <- hashed.model.matrix(c("habitat", "population"), mushroom, hash.size = 10,
                             create.mapping = TRUE)
mapping <- hash.mapping(mush)
mean(duplicated(mapping))

bst <- xgboost(data = mush, label = mushroom$class, nround = 10, params = list(
  booster = "gbtree",
  objective = "binary:logistic",
  eval_metric = "auc"
))
```

```{r}
# type_mush <- as.factor(c(mushroom$.))
# unclass(type_mush)
mushroom_df <- data.frame(mushroom)
mushroom_df$class <- as.factor(mushroom$class)
for (name_ in colnames(mushroom_df)[-1]){
     type_mush <- as.factor(c(mushroom[[name_]]))
     mushroom_df[name_] <- unclass(type_mush)
}

```

```{r}
# mushroom_df <- data.frame(mushroom)
# mush_features_lst <- list()
# i = 0
# for (name_ in colnames(mushroom_df)[-1]){
#     mush_features_lst[name_] <- name_
#     i<- i+1
# }
```


```{r}
# set.seed(101) # Set Seed so that same sample can be reproduced in future also
# Now Selecting 75% of data as sample from total 'n' rows of the data  
sample <- sample.int(n = nrow(mushroom_df), size = floor(.75*nrow(mushroom_df)), replace = F)
train_mushroom <- mushroom_df[sample, ]
test_mushroom  <- mushroom_df[-sample, ]
# (l <- sapply(iris, function(x) is.factor(x)))
```

```{r}
library(randomForest)
# set.seed(69)
rf_model<-rand_forest(trees = 25, mtry = 4, mode = "classification") %>%
#rf_model<-randomForest(trees = 25, mtry = 4, mode = "classification") %>% 
  set_engine("randomForest",
# importance = T to have permutation score calculated
  importance=TRUE,
# localImp=T for randomForestExplainer(next post)
  localImp = T, ) 
#%>% 
 # translate()
rf_model
```


```{r}
rf_mushroom = rf_model %>% 
  fit(class~., data=train_mushroom)
rf_mushroom
```

generate the predicted value
generate the  confusion matrix to evaluted the model
```{r}
pred  = predict(rf_mushroom, test_mushroom)
library(caret)
c_ <- confusionMatrix(pred$.pred_class,as.factor(test_mushroom$class))
c_
```


```{r}
tree_opt <- function(data, splt, y){
  # engine 
  engine <- c('C5.0','classification')
  
  # formula 
  form <- reformulate(termlabels = c('.'), response = c(y))
  # training split
  # train <- sample(1:nrow(data), nrow(data)*splt)
  # sample <- sample.int(n = nrow(data), size = floor(splt*nrow(data)), replace = F)
  train <- sample.int(n = nrow(data), size = floor(splt*nrow(data)), replace = F)
  # train <- data[sample, ]
  # test  <- data[-sample, ]
  # Bagger
  ctrl <- control_bag(var_imp = TRUE)
  bagg <- bagger(form, data = data[train,], base_model = engine[1])
  # boosted tree model 
  bt_model <-
    boost_tree(
      learn_rate = 0.3,
      trees = 5,
      tree_depth = 6,
      min_n = 1,
      sample_size = 1,
      mode = engine[2]
    ) %>% set_engine("xgboost", verbose = 2) %>%  fit(form, data = data[train,])
  # randomForest
  rf_model <-
  rand_forest(trees = 10, mtry = 5, mode = engine[2]) %>% 
    set_engine("randomForest",
    # importance = T to have permutation score calculated
    importance = T,
    # localImp=T for randomForestExplainer(next post)
    localImp = T,) %>% 
    fit(form, data = data[train,])
  # predictions 
  info.bag <- predict(bagg, data[-train,])
  info.bt <- predict(bt_model, data[-train,])
  info.rf <- predict(rf_model, data[-train,])
  # Metrics

  acc.bag <- confusionMatrix(info.bag$.pred_class, as.factor(data[-train,][[y]]))
  acc.bt <- confusionMatrix(info.bt$.pred_class, as.factor(data[-train,][[y]]))
  acc.rf <- confusionMatrix(info.rf$.pred_class, as.factor(data[-train,][[y]]))
  accuracy <- c(acc.bag$overall[1], acc.bt$overall[1], acc.rf$overall[1])

  # Result
  
  model <- c('Bagger', 'Boosted', 'RandomForest')
  data.frame(model, accuracy)
  #conf_matrix <- data.frame(Bagger=acc.bag, Boosted=acc.bt, RandomForest=acc.rf)

  return(list(Bagger=acc.bag, Boosted=acc.bt, RandomForest=acc.rf,
              Bag_mod=bagg, Boos_mod=bt_model, Rf_mod=rf_model))
}
```

```{r}
t <- tree_opt(data = mushroom_df, splt = 0.80, y ='class')
```

